---
title: "PCA: IBS Versions"
author: "Ryan Peek"
date: "*Updated: `r format(Sys.Date())`*"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = normalizePath("../"))

suppressPackageStartupMessages({
  library(tidyverse); 
  library(viridis); 
  library(plotly);
  library(ggforce);
  library(ggthemes)
  #library(patchwork)
})

```

# SNPs

The IBS angsd method using all sites, so doesn't necessarily call a single SNP. 

## Sierra Sites 

Looking at the Sierra Sites (this does not include Feather Basin which is Sierra/Basin Range)

### Sierra 100k

```{r sierra1PCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Sierra v1: PCA"}

reads <- "100k"
site <- "sierra"
# a few other options: sierra_basin_100k (this is only RABO but includes FEA)
# sierra_100k/50k (this is all sierras)
# sierra_v2 (this is only minus stan and tuo)
# sierra_v3 (this is the reg/unreg run minus deer crk, sfa, and tuo/stan)

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Filter outliers:
to_rm <- c(which(annot$SampleID=="RAP-122"), # outlier SFY Spring (166),
           which(annot$SampleID=="RAP-278"))#, # outlier Bear Missouri (13)
           #which(grepl("^DEER-|^SFA-",annot$Locality))) # SFA/DEER CK samples
covar<- covar[-c(to_rm),-c(to_rm)]
annot <- annot[-c(to_rm),]

# fix the Upper Feather Slate Creek Sample to Yuba:
#which(annot$SampleID=="RAP1716")
annot$HU_8_NAME[18] <- "Upper Yuba"
#annot$HU_8_NAME[which(annot$SampleID=="RAP1716")]

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)


# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

## GGPLOTLY
ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))
)

```

```{r sierra1PCAplot, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Sierra v1 static: PCA"}
# static
library(ggforce)
library(ggthemes)
ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2),
                           color=quote(HU_8_NAME),
                           text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +

  geom_point(size=4, alpha=0.8) +
  ggforce::geom_mark_circle(aes(group=HU_8_NAME), show.legend = F)+
  theme_bw(base_family = "Roboto Condensed") +
  #scale_color_grey("Locality") +
  #scale_color_colorblind() +
  scale_color_viridis_d("Locality", option = "D") +
  scale_shape_discrete("Locality") +
  ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))

#ggsave(filename = "figs/pca_fea_thresh_locality_colorblind.png", width = 8, height = 5, units = "in", dpi=300)

#ggsave(filename = "figs/pca_fea_thresh_locality_virD.png", width = 8, height = 5, units = "in", dpi=300)

```

### Sierra 100k v3

This is the reg/unreg sites only. It excludes Deer Creek, SFA, TUO, and STAN samples.

```{r sierra3PCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Sierra v3: PCA"}

reads <- "100k"
site <- "sierra_v2"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Filter outliers:
to_rm <- c(which(annot$SampleID=="RAP-122"), # outlier SFY Spring (166),
           which(annot$SampleID=="RAP-278"))#, # outlier Bear Missouri (13)
           #which(grepl("^DEER-|^SFA-",annot$Locality))) # SFA/DEER CK samples
covar<- covar[-c(to_rm),-c(to_rm)]
annot <- annot[-c(to_rm),]

# fix the Upper Feather Slate Creek Sample to Yuba:
slck <- which(annot$SampleID=="RAP1716")
annot$HU_8_NAME[slck] <- "Upper Yuba"
annot$HU_8_NAME[which(annot$SampleID=="RAP1716")]

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2

# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

# PLOT
ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))
)

```

### Sierra 50k

```{r sierra50PCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Sierra v1: PCA"}

reads <- "50k"
site <- "sierra"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_",reads, "_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

# wait what!!?
annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(Seq, .keep_all = T)

# Filter outliers:
to_rm <- c(which(annot$SampleID=="RAP-122"), # outlier SFY Spring (166),
           which(annot$SampleID=="RAP-278"))#, # outlier Bear Missouri (13)
           #which(grepl("^DEER-|^SFA-",annot$Locality))) # SFA/DEER CK samples
covar<- covar[-c(to_rm),-c(to_rm)]
annot <- annot[-c(to_rm),]

# fix the Upper Feather Slate Creek Sample to Yuba:
slck <- which(annot$SampleID=="RAP1716")
annot$HU_8_NAME[slck] <- "Upper Yuba"
annot$HU_8_NAME[which(annot$SampleID=="RAP1716")]


# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 3
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

# PLOT
ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))
)

```

### Sierra & Basin Range 100k

```{r sierrabasinPCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Sierra/Basin Range PCA"}

reads <- "100k"
site <- "sierra_basin"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Filter outliers:
to_rm <- c(which(annot$SampleID=="RAP-122"), # outlier SFY Spring (166),
           which(annot$SampleID=="RAP-278"))#, # outlier Bear Missouri (13)
           #which(grepl("^DEER-|^SFA-",annot$Locality))) # SFA/DEER CK samples
covar<- covar[-c(to_rm),-c(to_rm)]
annot <- annot[-c(to_rm),]

# fix the Upper Feather Slate Creek Sample to Yuba:
slck <- which(annot$SampleID=="RAP1716")
annot$HU_8_NAME[slck] <- "Upper Yuba"
annot$HU_8_NAME[which(annot$SampleID=="RAP1716")]

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

# plot
ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("Sierra/Basin RABO: ", reads, " [",title,"]"))
)

```

```{r staticplotsierrabasin100k, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Sierra/Basin Range PCA static"}

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

# static
library(ggforce)
library(ggthemes)
ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2),
                           color=quote(HU_8_NAME),
                           text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +

  geom_point(size=4, alpha=0.8) +
  ggforce::geom_mark_circle(aes(group=HU_8_NAME), show.legend = F)+
  theme_bw(base_family = "Roboto Condensed") +
  #scale_color_grey("Locality") +
  #scale_color_colorblind() +
  scale_color_viridis_d("Locality", option = "D") +
  scale_shape_discrete("Locality") +
  ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))

# PC's:
pc1 <- 1
pc2 <- 3
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

# static
ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2),
                           color=quote(HU_8_NAME),
                           text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +

  geom_point(size=4, alpha=0.8) +
  ggforce::geom_mark_circle(aes(group=HU_8_NAME), show.legend = F)+
  theme_bw(base_family = "Roboto Condensed") +
  #scale_color_grey("Locality") +
  #scale_color_colorblind() +
  scale_color_viridis_d("Locality", option = "D") +
  scale_shape_discrete("Locality") +
  ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))

# PC's:
pc1 <- 2
pc2 <- 3
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

# static
ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2),
                           color=quote(HU_8_NAME),
                           text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +

  geom_point(size=4, alpha=0.8) +
  ggforce::geom_mark_circle(aes(group=HU_8_NAME), show.legend = F)+
  theme_bw(base_family = "Roboto Condensed") +
  #scale_color_grey("Locality") +
  #scale_color_colorblind() +
  scale_color_viridis_d("Locality", option = "D") +
  scale_shape_discrete("Locality") +
  ggtitle(paste0("Sierra RABO: ", reads, " [",title,"]"))

```

## NFA Watershed 100k

```{r nfaPCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="MFA 100k: PCA"}

reads <- "100k"
site <- "nfa"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)


# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

#plot
ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("NFA RABO: ", reads, " [",title,"]"))
)
```

## MFA Watershed 100k

```{r mfaPCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="MFA 100k: PCA"}

reads <- "100k"
site <- "mfa"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)


# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)


# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

#plot
ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("MFA RABO: ", reads, " [",title,"]"))
)
```

## MFA Hydro Sites

Looking at the MFA Sites in hydropeaking reach only at 100k.

### MFA Hydro 100k v1

```{r mfahydro1PCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="MFA Hydro v1: PCA"}

reads <- "100k"
site <- "mfa_hydro"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# read covar file
#covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k.covar")
#covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)


# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)


# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")
# title

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("MFA Hydro: ", reads, " [",title,"]"))
)
```

```{r mfahydro1PCAplot, warning=FALSE, echo=F, eval=F, message=FALSE, fig.cap="MFA Hydro v1: PCA static"}

# static
library(ggforce)
library(ggthemes)
ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2),
                           color=quote(River), shape=quote(HU_12_NAME),
                           text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +

  geom_point(size=4, alpha=0.8) +
  ggforce::geom_mark_circle(aes(group=River), show.legend = F)+
  theme_bw(base_family = "Roboto Condensed") +
  #scale_color_grey("Locality") +
  #scale_color_colorblind() +
  scale_color_viridis_d("Locality", option = "D") +
  scale_shape_discrete("Locality") +
  ggtitle(paste0("MFA Hydro: ", reads, " [",title,"]"))

#ggsave(filename = "figs/pca_fea_thresh_locality_colorblind.png", width = 8, height = 5, units = "in", dpi=300)

#ggsave(filename = "figs/pca_fea_thresh_locality_virD.png", width = 8, height = 5, units = "in", dpi=300)

```

### MFA Hydro 100k v2

```{r mfahydro2PCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="MFA Hydro v2: PCA"}

reads <- "100k"
site <- "mfa_hydro2"
siteclst <- "mfa_hydro"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# read covar file
#covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k.covar")
#covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",siteclst,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)


# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)


# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")
# title

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("MFA Hydro: ", reads, " [",title,"]"))
)

```

## Feather Watershed 100k

```{r feaPCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Feather 100k: PCA"}

reads <- "100k"
site <- "fea"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)


# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)


# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")
# title

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size = 9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("Feather RABO: ", reads, " [",title,"]"))
)

```

## Yuba

### Middle Yuba Watershed 100k

```{r mfyPCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="MFY 100k: PCA"}

reads <- "100k"
site <- "mfy"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("MFY RABO: ", reads, " [",title,"]"))
)
```

### Yuba 100k v1

```{r yubPCAplotly, warning=FALSE, echo=FALSE, eval=T, message=FALSE, fig.cap="Yuba 100k: PCA"}

reads <- "100k"
site <- "yuba"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",site,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)


# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("Yuba RABO: ", reads, " [",title,"]"))
)
```


### Yuba 100k v2

```{r yub2PCAplotly, warning=FALSE, echo=FALSE, eval=F, message=FALSE, fig.cap="Yuba v2: PCA"}

reads <- "100k"
site <- "yuba2"
siteclst <- "yuba"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",siteclst,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2

# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("RABO: ", reads, " [",title,"]"))
)

```




## Feather

### Feather 100k

```{r feaPCAplotly100k, warning=FALSE, echo=FALSE, eval=F, message=FALSE, fig.cap="MFY 100k: PCA"}

reads <- "100k"
site <- "fea"
siteclst <- "fea"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",siteclst,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("FEA RABO: ", reads, " [",title,"]"))
)
```


```{r fea2PCAplotly, warning=FALSE, echo=FALSE, eval=F, message=FALSE, fig.cap="FEA2 100k: PCA"}

reads <- "100k"
site <- "fea2"
siteclst <- "fea"

# read covar file
covar_path <- paste0("./data_output/angsd/rabo_",site, "_100k_thresh.covMat")
covar <- read.table(covar_path, stringsAsFactors = F)

# Read annot file
annot <- read.table(paste0("data_output/bamlists/bamlist_mrg_RABO_",siteclst,"_",reads,
                           "_clst"),stringsAsFactors = F, header = TRUE)

# combine with metadata
suppressMessages(metadat<- read_csv("data_output/rapture06_metadata_revised.csv"))

annot <- inner_join(metadat, annot, by=c("SampleID"="IID")) %>% 
  arrange(Seq) %>% 
  distinct(SampleID, .keep_all = T)

# Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
PC$River <- factor(annot$Locality)
PC$Pop <- factor(annot$HU_12_NAME) # CLUSTER
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_10_NAME <- factor(annot$HU_10_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)

# PC's:
pc1 <- 1
pc2 <- 2
# Title (and % explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

ggplotly(p = 
           ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2), color=quote(River), text=quote((paste0("ID: ", ID, " <br> River: ", River))))) +
           geom_point(size=4, alpha=0.8) + 
           theme_bw(base_size=9) +
           scale_color_viridis_d() + 
           ggtitle(paste0("FEA2 RABO: ", reads, " [",title,"]"))
)
```